\vskip-0.6in \Large Political Uncertainty

### 2016-02-09 two-groups.tex
move intuition section and change subsecs to full sections

@@

\newpage		
\subsection{Second Order Conditions}
Perhaps more importantly, let's look at the second order condition (using the same substitutions at the top of the previous page). The first derivative of the objective function w.r.t. $b(-\frac{1}{2})$ is (in logistic form)
\begin{equation}
  W_B\frac{e^{-Z}}{\left(1+e^{-Z}\right)^2}\left[\frac{1}{1+e^{-X}} + \frac{1}{1+e^{-Y}} - 2\frac{1}{1+e^{-X}}  \frac{1}{1+e^{-Y}} \right] - 1
\end{equation}
\begin{equation}
  W_B\frac{e^{-Z}}{\left(1+e^{-Z}\right)^2}\left[\frac{e^{-X} + e^{-Y}}{\left(1+e^{-X}\right)\left(1+e^{-Y}\right)} \right] - 1
	\label{eq:firstderiv}
\end{equation}
Now, the second derivative:
\[
  W_B\left[\frac{e^{-X} + e^{-Y}}{\left(1+e^{-X}\right)\left(1+e^{-Y}\right)} \right] \left[ \frac{e^{-Z} 2\left(1+e^{-Z}\right) e^{-Z}-\left(1+e^{-Z}\right)^2 e^{-Z}}{\left(1+e^{-Z}\right)^4} \right]
\]
\[
  W_B\left[\frac{e^{-X} + e^{-Y}}{\left(1+e^{-X}\right)\left(1+e^{-Y}\right)} \right] \left[ \frac{e^{-Z} 2 e^{-Z}-\left(1+e^{-Z}\right) e^{-Z}}{\left(1+e^{-Z}\right)^3} \right]
\]
\[
  W_B\left[\frac{e^{-X} + e^{-Y}}{\left(1+e^{-X}\right)\left(1+e^{-Y}\right)} \right] \left[ \frac{e^{-Z} \left\{2 e^{-Z}-\left(1+e^{-Z}\right) \right\}}{\left(1+e^{-Z}\right)^3} \right]
\]
\[
  W_B\left[\frac{e^{-X} + e^{-Y}}{\left(1+e^{-X}\right)\left(1+e^{-Y}\right)} \right] \left[ \frac{e^{-Z} \left\{2 e^{-Z}-1-e^{-Z} \right\}}{\left(1+e^{-Z}\right)^3} \right]
\]
\[
  W_B\left[\frac{e^{-X} + e^{-Y}}{\left(1+e^{-X}\right)\left(1+e^{-Y}\right)} \right] \left[ \frac{e^{-Z} \left\{e^{-Z}-1 \right\}}{\left(1+e^{-Z}\right)^3} \right]
\]
\[
  W_B\left[\frac{e^{-X} + e^{-Y}}{\left(1+e^{-X}\right)\left(1+e^{-Y}\right)} \right] \left[ \frac{e^{-Z}}{\left(1+e^{-Z}\right)^3} \right] \left\{e^{-Z}-1 \right\}
\]
Everything except for the expression in the curly braces is positive. The expression in the curly braces is positive when $z\leq0$ and negative when $z\geq0$. So this objective function appears to have an inflection point at $z=0$.
\begin{itemize}
	\item When $z < 0$ (as when this legislator is naturally against you and you don't bribe her), this expression is negative
\end{itemize}

@@

\newpage
\subsection{Intuition}
\begin{itemize}
	\item A legislator votes for $x$ (A's preferred policy, against the status quo) if 
		\[
		  v(z) = \al - \beta z + a(z) - b(z) > 0
		\]
	\item B needs 2 ``no'' votes, or two legislators with $v(z) < 0$. When $a(z) = 0$, this is
		\[
		  \al - \beta z - b(z) < 0
		\]

\newpage
\subsection{One Non-negative Bribe}
When $\al = 0$ and $WB=8$, we find that $b\left(-\frac{1}{2}\right)$ and $b\left(\frac{1}{2}\right)$ are zero. Only $b(0)$ is positive.
\begin{itemize}
	\item Only the FOC for $X$ holds with equality (i.e. $\lambda_X = 0$)
		\[
			\frac{e^{-Y} + e^{-Z}}{\left(1+e^{-Y}\right)\left(1+e^{-Z}\right)} \frac{e^{-X}}{\left(1+e^{-X}\right)^2}= \frac{1}{W_B}
		\]
	\item We know exactly what $Y$ and $Z$ will be since $b\left(-\frac{1}{2}\right)$ and $b\left(\frac{1}{2}\right)$ are zero. Given $WB$, we can solve for $X$ and then $b(0)$.
	\item Looking at the SOC, when there are three inequality constraints and 2 bind, we only need a condition on the last principal minor of a suitably constructed bordered Hessian.
		\begin{itemize}
			\item It should be negative because $n=3$
			\item It's a 5x5 matrix because there are two binding inequality constraints.
		\end{itemize}
		The SOCs when one non-negativity constraint binds; bordered Hessian is 5x5. \\
		\begin{tabular}{ccccc}
			0 & 0 & $\frac{\partial g_1}{\partial b\left(-\frac{1}{2}\right)}$ & $\frac{\partial g_1}{\partial b\left(-0\right)}$ & $\frac{\partial g_1}{\partial b\left(\frac{1}{2}\right)}$ \\
			0 & 0 & $\frac{\partial g_3}{\partial b\left(-\frac{1}{2}\right)}$ & $\frac{\partial g_3}{\partial b\left(-0\right)}$ & $\frac{\partial g_3}{\partial b\left(\frac{1}{2}\right)}$ \\
			$\frac{\partial g_1}{\partial b\left(-\frac{1}{2}\right)}$ & $\frac{\partial g_3}{\partial b\left(-\frac{1}{2}\right)}$& $\frac{\partial^2 L}{\partial b\left(-\frac{1}{2}\right)^2}$ & $\frac{\partial^2 L}{\partial b(0)\partial b\left(-\frac{1}{2}\right)}$ & $\frac{\partial^2 L}{\partial b\left(\frac{1}{2}\right)\partial b\left(-\frac{1}{2}\right)}$  \\
			$\frac{\partial g_1}{\partial b\left(0\right)}$ & $\frac{\partial g_3}{\partial b\left(-0\right)}$ & $\frac{\partial^2 L}{\partial b\left(-\frac{1}{2}\right)\partial b\left(0\right)}$ & $\frac{\partial^2 L}{\partial b(0)^2}$ & $\frac{\partial^2 L}{\partial b\left(\frac{1}{2}\right)\partial b\left(0\right)}$ \\
			$\frac{\partial g_1}{\partial b\left(\frac{1}{2}\right)}$ & $\frac{\partial g_3}{\partial b\left(\frac{1}{2}\right)}$ & $\frac{\partial^2 L}{\partial b\left(-\frac{1}{2}\right)\partial b\left(\frac{1}{2}\right)}$ & $\frac{\partial^2 L}{\partial b(0) \partial b\left(\frac{1}{2}\right)}$ & $\frac{\partial^2 L}{\partial b\left(\frac{1}{2}\right)^2}$
		\end{tabular} \\

@@

\newpage
\subsection{All Three Bribes are Non-negative}
In many cases all three bribes are non-negative. Here are very loose results
\begin{itemize}
	\item I always observe that $X=Y=Z$. I think it's easy to show this analytically.
	\item From the FOCs, when $X$ and $Y$ are non-negative, it can be shown that a set of inequalities just below Equations~\ref{eq:1} and \ref{eq:2} must hold.
	\item The SOCs when 0 non-negativity constraints bind are on the 3x3 matrix of second derivatives. There are three conditions:
		\[
		  \frac{\partial^2 L}{\partial b\left(-\frac{1}{2}\right)^2} \left\{\frac{\partial^2 L}{\partial b(0)^2} \frac{\partial^2 L}{\partial b\left(\frac{1}{2}\right)^2} - \left[\frac{\partial^2 L}{\partial b\left(\frac{1}{2}\right)\partial b\left(0\right)}\right]^2 \right\} \leq 0
		\]
		\[
		  \frac{\partial^2 L}{\partial b(0)^2} \frac{\partial^2 L}{\partial b\left(\frac{1}{2}\right)^2} - \left[\frac{\partial^2 L}{\partial b\left(\frac{1}{2}\right)\partial b\left(0\right)}\right]^2 \geq 0
		\]
		\[
		  \frac{\partial^2 L}{\partial b\left(\frac{1}{2}\right)^2} \leq 0
		\]
		The last one says that $Y$ must be non-negative. The first and second together say that $Z$ must be non-negative.
	\item $Y$ non-negative combined with the inequalities below Equations ~\ref{eq:1} and \ref{eq:2} tells us that $X=Y > 0$.
	\item Similar logic that leads to those inequalities can be used to derive analagous inequalities for $X$ and $Z$ (or $Y$ and $Z$). Then $Z$ non-negative tells us $Z=Y>0$ so that $X=Y=Z >0$.
\end{itemize}

@@

\newpage
\subsection{Numerical Results}	
For all of these, $\beta = 1$ and there are three legislators with $b\in\left\{-\frac{1}{2},0,\frac{1}{2}\right\}$.
\begin{itemize}
	\item Below, the column headings are $v(z)$ before the bribe.
	\item That is, from left to right, $-Z,-Y,-X$.
\end{itemize}

>>>>

\newpage
\section{Intuition}
\begin{itemize}
	\item A legislator votes for $x$ (A's preferred policy, against the status quo) if 
		\[
		  v(z) = \al - \beta z + a(z) - b(z) > 0
		\]
	\item B needs 2 ``no'' votes, or two legislators with $v(z) < 0$. When $a(z) = 0$, this is
		\[
		  \al - \beta z - b(z) < 0
		\]

@@

\newpage		
\section{Second Order Conditions}
Perhaps more importantly, let's look at the second order condition (using the same substitutions at the top of the previous page). The first derivative of the objective function w.r.t. $b(-\frac{1}{2})$ is (in logistic form)
\begin{equation}
  W_B\frac{e^{-Z}}{\left(1+e^{-Z}\right)^2}\left[\frac{1}{1+e^{-X}} + \frac{1}{1+e^{-Y}} - 2\frac{1}{1+e^{-X}}  \frac{1}{1+e^{-Y}} \right] - 1
\end{equation}
\begin{equation}
  W_B\frac{e^{-Z}}{\left(1+e^{-Z}\right)^2}\left[\frac{e^{-X} + e^{-Y}}{\left(1+e^{-X}\right)\left(1+e^{-Y}\right)} \right] - 1
	\label{eq:firstderiv}
\end{equation}
Now, the second derivative:
\[
  W_B\left[\frac{e^{-X} + e^{-Y}}{\left(1+e^{-X}\right)\left(1+e^{-Y}\right)} \right] \left[ \frac{e^{-Z} 2\left(1+e^{-Z}\right) e^{-Z}-\left(1+e^{-Z}\right)^2 e^{-Z}}{\left(1+e^{-Z}\right)^4} \right]
\]
\[
  W_B\left[\frac{e^{-X} + e^{-Y}}{\left(1+e^{-X}\right)\left(1+e^{-Y}\right)} \right] \left[ \frac{e^{-Z} 2 e^{-Z}-\left(1+e^{-Z}\right) e^{-Z}}{\left(1+e^{-Z}\right)^3} \right]
\]
\[
  W_B\left[\frac{e^{-X} + e^{-Y}}{\left(1+e^{-X}\right)\left(1+e^{-Y}\right)} \right] \left[ \frac{e^{-Z} \left\{2 e^{-Z}-\left(1+e^{-Z}\right) \right\}}{\left(1+e^{-Z}\right)^3} \right]
\]
\[
  W_B\left[\frac{e^{-X} + e^{-Y}}{\left(1+e^{-X}\right)\left(1+e^{-Y}\right)} \right] \left[ \frac{e^{-Z} \left\{2 e^{-Z}-1-e^{-Z} \right\}}{\left(1+e^{-Z}\right)^3} \right]
\]
\[
  W_B\left[\frac{e^{-X} + e^{-Y}}{\left(1+e^{-X}\right)\left(1+e^{-Y}\right)} \right] \left[ \frac{e^{-Z} \left\{e^{-Z}-1 \right\}}{\left(1+e^{-Z}\right)^3} \right]
\]
\[
  W_B\left[\frac{e^{-X} + e^{-Y}}{\left(1+e^{-X}\right)\left(1+e^{-Y}\right)} \right] \left[ \frac{e^{-Z}}{\left(1+e^{-Z}\right)^3} \right] \left\{e^{-Z}-1 \right\}
\]
Everything except for the expression in the curly braces is positive. The expression in the curly braces is positive when $z\leq0$ and negative when $z\geq0$. So this objective function appears to have an inflection point at $z=0$.
\begin{itemize}
	\item When $z < 0$ (as when this legislator is naturally against you and you don't bribe her), this expression is negative
\end{itemize}

@@

\newpage
\section{One Non-negative Bribe}
When $\al = 0$ and $WB=8$, we find that $b\left(-\frac{1}{2}\right)$ and $b\left(\frac{1}{2}\right)$ are zero. Only $b(0)$ is positive.
\begin{itemize}
	\item Only the FOC for $X$ holds with equality (i.e. $\lambda_X = 0$)
		\[
			\frac{e^{-Y} + e^{-Z}}{\left(1+e^{-Y}\right)\left(1+e^{-Z}\right)} \frac{e^{-X}}{\left(1+e^{-X}\right)^2}= \frac{1}{W_B}
		\]
	\item We know exactly what $Y$ and $Z$ will be since $b\left(-\frac{1}{2}\right)$ and $b\left(\frac{1}{2}\right)$ are zero. Given $WB$, we can solve for $X$ and then $b(0)$.
	\item Looking at the SOC, when there are three inequality constraints and 2 bind, we only need a condition on the last principal minor of a suitably constructed bordered Hessian.
		\begin{itemize}
			\item It should be negative because $n=3$
			\item It's a 5x5 matrix because there are two binding inequality constraints.
		\end{itemize}
		The SOCs when one non-negativity constraint binds; bordered Hessian is 5x5. \\
		\begin{tabular}{ccccc}
			0 & 0 & $\frac{\partial g_1}{\partial b\left(-\frac{1}{2}\right)}$ & $\frac{\partial g_1}{\partial b\left(-0\right)}$ & $\frac{\partial g_1}{\partial b\left(\frac{1}{2}\right)}$ \\
			0 & 0 & $\frac{\partial g_3}{\partial b\left(-\frac{1}{2}\right)}$ & $\frac{\partial g_3}{\partial b\left(-0\right)}$ & $\frac{\partial g_3}{\partial b\left(\frac{1}{2}\right)}$ \\
			$\frac{\partial g_1}{\partial b\left(-\frac{1}{2}\right)}$ & $\frac{\partial g_3}{\partial b\left(-\frac{1}{2}\right)}$& $\frac{\partial^2 L}{\partial b\left(-\frac{1}{2}\right)^2}$ & $\frac{\partial^2 L}{\partial b(0)\partial b\left(-\frac{1}{2}\right)}$ & $\frac{\partial^2 L}{\partial b\left(\frac{1}{2}\right)\partial b\left(-\frac{1}{2}\right)}$  \\
			$\frac{\partial g_1}{\partial b\left(0\right)}$ & $\frac{\partial g_3}{\partial b\left(-0\right)}$ & $\frac{\partial^2 L}{\partial b\left(-\frac{1}{2}\right)\partial b\left(0\right)}$ & $\frac{\partial^2 L}{\partial b(0)^2}$ & $\frac{\partial^2 L}{\partial b\left(\frac{1}{2}\right)\partial b\left(0\right)}$ \\
			$\frac{\partial g_1}{\partial b\left(\frac{1}{2}\right)}$ & $\frac{\partial g_3}{\partial b\left(\frac{1}{2}\right)}$ & $\frac{\partial^2 L}{\partial b\left(-\frac{1}{2}\right)\partial b\left(\frac{1}{2}\right)}$ & $\frac{\partial^2 L}{\partial b(0) \partial b\left(\frac{1}{2}\right)}$ & $\frac{\partial^2 L}{\partial b\left(\frac{1}{2}\right)^2}$
		\end{tabular} \\

@@

\newpage
\section{All Three Bribes are Non-negative}
In many cases all three bribes are non-negative. Here are very loose results
\begin{itemize}
	\item I always observe that $X=Y=Z$. I think it's easy to show this analytically.
	\item From the FOCs, when $X$ and $Y$ are non-negative, it can be shown that a set of inequalities just below Equations~\ref{eq:1} and \ref{eq:2} must hold.
	\item The SOCs when 0 non-negativity constraints bind are on the 3x3 matrix of second derivatives. There are three conditions:
		\[
		  \frac{\partial^2 L}{\partial b\left(-\frac{1}{2}\right)^2} \left\{\frac{\partial^2 L}{\partial b(0)^2} \frac{\partial^2 L}{\partial b\left(\frac{1}{2}\right)^2} - \left[\frac{\partial^2 L}{\partial b\left(\frac{1}{2}\right)\partial b\left(0\right)}\right]^2 \right\} \leq 0
		\]
		\[
		  \frac{\partial^2 L}{\partial b(0)^2} \frac{\partial^2 L}{\partial b\left(\frac{1}{2}\right)^2} - \left[\frac{\partial^2 L}{\partial b\left(\frac{1}{2}\right)\partial b\left(0\right)}\right]^2 \geq 0
		\]
		\[
		  \frac{\partial^2 L}{\partial b\left(\frac{1}{2}\right)^2} \leq 0
		\]
		The last one says that $Y$ must be non-negative. The first and second together say that $Z$ must be non-negative.
	\item $Y$ non-negative combined with the inequalities below Equations ~\ref{eq:1} and \ref{eq:2} tells us that $X=Y > 0$.
	\item Similar logic that leads to those inequalities can be used to derive analagous inequalities for $X$ and $Z$ (or $Y$ and $Z$). Then $Z$ non-negative tells us $Z=Y>0$ so that $X=Y=Z >0$.
\end{itemize}

@@

\newpage
\section{Numerical Results}	
For all of these, $\beta = 1$ and there are three legislators with $b\in\left\{-\frac{1}{2},0,\frac{1}{2}\right\}$.
\begin{itemize}
	\item Below, the column headings are $v(z)$ before the bribe.
	\item That is, from left to right, $-Z,-Y,-X$.
\end{itemize}
